### 2012-03-01 3.tex
Drastic restructuring and some writing on chapters 2-3.

% 3: Static and derived constraints on Turkish vowel harmony

In the preceding chapter, it was shown that apparent underrepresentation of certain sequences in the lexicon may occur

This brief chapter argues that even statistically valid phonotactic generalizations are ignroed by speakers. Two phonotactic generalizations over the Turksih lexicon are shown to be equally statistically reliable, but only the generalization which is derived by a phonological alternation influence wordlikeness judgements.

Research on gradient phonotactics has usually rejected this possibilty implicit
ly.

After reviewing co-occurrence restrictions in Gitksan roots, \citet{Brown2010} rejects this explicitly. He writes:

@@

@@

%\ex Arabic verbs of coming into being: \\
%\begin{tabular}{r l l l}
%a. & kabura & `become big'       & (cf. \emph{kabiːr} `big') \\
%b. & saʁiir & `become small'     & (cf. \emph{saʁiːr} `small') \\
%c. & ħasuna & `become beautiful' & (cf. \emph{ħasan} `handsome') \\
%\end{tabular}
%\xe

>>>>

\chapter{Static and derived constraints in Turkish}
\label{turkish}

One of the original motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps. \citet{Chomsky1965} propose that a theory of phonology must account for speakers' abilities to distinguish between accidental and structural gaps in their language, citing a contrast between \emph{blick}, which is judged to be a possible but unattested word of English, and \emph{bnick}, which is judged to be impossible. \emph{SPE} (p.~3810f.) places this contrast in a system of a system of sequence structure rules that also are used to eliminate lexical redundancy. In the phonological theory of the era, there is no process that could generate the \emph{blick} $\sim$ \emph{bnick} contrast or numerous other lexical regularities \citep[][283f.]{Anderson1974}.\footnote{This sequence of arguments has already been critiqued in Chapter \ref{1}, where it was shown that prosodic licensing and Stampean occultation places this contrast in core phonology, and that the principles by which speakers recognize impossible words must be stated on surface representations, not underlying representations. Further, \citet[][528f.]{Halle1975} rejects his own principle that lexical entries should be free of all redundancies (see \citealt[][201]{Reiss2003a} and \citealt{Vaux2003} for further discussion). This rejected principle of a redundancy-free lexicon is one of the two principles that motivates Morpheme Structure Constraints; the other is the subject of this and the following chapter.}

the other being the subject

a principle which is one of the two motivations  motivated the introduction of morpheme structure constraints.

In the preceding chapter, it was shown that apparent underrepresentation of certain sequences in the lexicon does not always require static constraints on possible underlying representations, since such distributions arise by chance and static constraints defined over natural classes contribute little to predicting the contents of the lexicon. 
%This result does not directly bear on the question of whether static restrictions have any synchronic reality.

Yet countless static constraints are statistically supported. Under the null hypothesis advanced in Chapter \ref{MSCs}, the lexical statistics of co-occurrence restrictions are irrelevant: the language acquisition device only attends to derived restrictions and ignores static restrictions regardless of their statistical reliability. On the other hand, many linguists assume that statistical significance provides sufficient conditions for positing synchronically active static constraints. For instance, \citet{Brown2010} writes:

@@

\noindent
A nuanced version of this alternative hypothesis would recognize that there may be other necessary conditions. Any explicit theory of static constraints will have to impose formal restrictions on possible co-occurrence restrictions. The set of possible constraints must respect results in formal learning theory, and may also be, for example, constrained by naturalness biases or formal markedness considerations.

A further widespread assumption is that static co-occurrence restrictions validated statistically will be reflected by wordlikeness judgements. This assumption is so deeply entrenched that a number of recent studies of phonotactic learning forgo wordlikeness judgements altogether and instead model the lexical statistics themselves (e.g., \citealt{Coetzee2008a} and \citealt{Anttila2008} on Muna and Arabic, \citealt[][385]{Hayes2008a} on Shona and Wargamay, \citealt{Brown2010} on Gitksan).

Since the null hypothesis does not countenance statistical significance as evidence for the synchronic reality of a static lexical pattern, such patterns do not adjudicate between the null hypothesis and the alternative, and another source of evidence is needed. In the remainder of Part I, I turn to data from wordlikeness judgement tasks for these purposes.

In a wordlikeness task, a speaker is presented with nonce words and asked to report such intuitions about the ``possibility'' of a word.

The alternative hypothesis, then, is that a statistically reliable co-occurrence restriction will be reflected in native speakers' wordlikeness judgements.

The data below, drawn from Turkish, shows that

This chapter takes the argument one step farther by showing a contrast between static and derived constraints on URs in Turkish. Whereas the derived constraint has a robust effect on native speakers' wordlikeness judgements, the static constraint, which is equally statistically valid, has no effect on these judgements. Anticipating the conclusion, the simplest explanation for the non-effect of this static constraint on wordlikeness judgements is that speakers do not internalize static co-occurrence restrictions at all.

@@

%It is important to note that the nature of the \emph{blick}/\emph{bnick} contrast is no longer a mystery. 
%Generative phonological theory at the time of \emph{SPE} lacked the representational vocabulary to directly distinguish between word-initial /bn/, which is disallowed, and /bn/ as a licit syllable contact cluster (e.g., \emph{o}[b.n]\emph{oxious}), a point first made by \citep{Hooper1973}. It is not the introduction of prosodic primitives into generative phonology that explains the contrast between \emph{blick} and \emph{bnick}, though, but rather the principle that surface forms must be syllabified \citep[e.g.,][63f.]{Kiparsky1982b} which \emph{bnick} fails. If, as is standardly assumed, phonology repairs unsyllabifiable outputs \citep[e.g.][]{Ito1989a,Noske1992}, then \emph{bnick} is an impossible output, doomed to be modified (perhaps to \emph{nick}; \citealt[][19f.]{Wolf2009}), and /bnɪk/ is an impossible UR by Stampean occultation.

\section{Turkish vowel MSCs} \input{3.1}
\section{Evaluation}         \input{3.2}
\section{Conclusion}         \input{3.3}

### 2012-03-01 4.2.1.tex
Drastic restructuring and some writing on chapters 2-3.

\subsubsection{Prediction distributions}

@@

\citet{Hayes2008a}
\citet{Albright2009}

@@

% EM
%\citet{EM}

@@

>>>>

\ex EM test results: \vspace{6pt} \\
\begin{tabular}{l | r r | r r r r r r | r r}
\toprule
     & \multicolumn{2}{c|}{one $\mathcal{N}$} & \multicolumn{6}{c|}{mixture of two $\mathcal{N}$s} & \multicolumn{2}{c}{log $\mathcal{L}$ test} \\
     & $\mu$ & $\sigma$ & $\mu_1$ & $\sigma_1$ & mix$_1$ & $\mu_2$ & $\sigma_2$ & mix$_2$ & $\Lambda$ & $p$-value  \\
\midrule
G\&J & 0.378 & 0.295    & 0.150   & 0.090      & 0.436     & 0.554   & 0.278    & 0.564     & 5.190    & 0.158   \\
S    & 0.596 & 0.361    & 0.236   & 0.232      & 0.437     & 0.876   & 0.105    & 0.563 & 48.965    & 1.3\e{-10} \\
A\&H & 0.431 & 0.266    & 0.301   & 0.200      & 0.559     & 0.595   & 0.246    & 0.411 & 16.186    & 0.001      \\
\bottomrule
\end{tabular}
\xe

%\ex Residualized correlations
%\begin{tabular}
%\toprule
%ASDF
%\midrule
%G\&J & 
%S    & 
%A\&H & 
%\bottomrule
%\xe

%\subsubsection{Prediction distributions}

@@

\citet{Hayes2008a}
\citet{Albright2009a}

@@

% EM
\citet{EM}

\footnote{In performing this evaluation, I have benefitted greatly from course notes by Mark Liberman and Stephen Isard.}
%available at \url{http://www.ling.upenn.edu/courses/cogs501/K-meansHW.html}

@@

where the difference in degrees o freedom equals $D$

%\ex $\displaystyle \textrm{Reject } H_0 \textrm{ iff } 
%\ex $\displaystyle \Lambda = -2 \textrm{ ln } L_0 + 2 \textrm{ ln } L_1}$ \xe
% > χ^2_{d.f.=D}$ \xe
